Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational

### 2016-09-19 ms_pre_post.Rmd
Make some minor wording tweeks and correct typos

---
title: 'Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational
  approach'
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation.
Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate phenotypes including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of inbreeding avoidance through fertilisation bias causes selection for evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that that inbreeding avoidance through mate choice but not fertilisation bias will evolve when inbreeding mating strategy is less costly than the combined cost of inbreeding fertilisation strategy and tendency for polyandry, while inbreeding avoidance through fertilisation bias but not mate choice will evolve when inbreeding mating strategy is more costly than the combined costs of inbreeding fertilisation strategy and tendency for polyandry. Finally, we test the hypothesis that constraints on evolution of one inbreeding strategy (e.g., inbreeding mating strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., inbreeding fertilisation strategy), and that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.

>>>>

---
title: 'Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational
  approach'
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation.

Here we apply a model of female and male reproductive interactions and associated inbreeding depression in offspring viability to quantify evolution of three separate phenotypes including female tendency for polyandry, female pre-copulatory inbreeding strategy (mating strategy), and female post-copulatory inbreeding strategy (fertilisation strategy). We use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of inbreeding avoidance through fertilisation bias causes selection for evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that inbreeding avoidance through mate choice but not fertilisation bias will evolve when inbreeding mating strategy is less costly than the combined cost of inbreeding fertilisation strategy and tendency for polyandry, while inbreeding avoidance through fertilisation bias but not mate choice will evolve when inbreeding mating strategy is more costly than the combined costs of inbreeding fertilisation strategy and tendency for polyandry. Finally, we test the hypothesis that constraints on evolution of one inbreeding strategy (e.g., inbreeding mating strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., inbreeding fertilisation strategy), and that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.
